---
title: "DATA_605_Assign_14_DThonn"
file: DATA_605_Assign_14_DThonn
name: Daniel Thonn
output:
  html_document: default
  pdf_document: default
course: DATA 605 CUNY MSDA program
---

** DATA_605_Assignment 14 - Calculus 2 - Taylor Series **


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``````{r, echo=TRUE}
if (!require(stats)) install.packages("stats")
library(stats)

if (!require(pracma)) install.packages("pracma")
library(pracma)


```


Assignment 14:

Work out the Taylor Series expansions of popular functions as shown below.

For each function, only consider its valid ranges as indicated in the notes when you are
computing the Taylor Series expansion.



** Question-1 **

1).
f(x) = 1/(1-x)

$$f(x) = \frac{1}{(1-x)}$$

Derivatives:
$$f(x) = \frac{1}{(1-x)}$$
$$f'(x) = \frac{1}{(1-x)^2}$$
$$f''(x) = \frac{2}{(1-x)^3}$$
$$f'''(x) = \frac{6}{(1-x)^4}$$
Taylor Series:
$$  = f(x) + \frac{f'(x)}{(1!)}(x-c)^1 + \frac{f'"(x)}{(2!)}(x-c)^2 + .... $$
$$  = f(x) + \frac{f'(x)}{(1!)}(x-c)^1 + \frac{f'"(x)}{(2!)}(x-c)^2 + .... $$
when plug in functions equals:
$$  = 1 + x + x^2 + x^3 + .... $$
Calculate in R:

``````{r, echo=TRUE}
# check range -10 to 10
a1 = -10
f1 <- function(x) {1/(1-x)} 
Result1 <- taylor(f1, x0=a1, 10)
Result1

plot(Result1)

# check range -5 to 5
a2 = -5
f2 <- function(x) {1/(1-x)} 
Result2 <- taylor(f2, x0=a2, 5)
Result2

plot(Result2)

# check range 3 to 5
a3 = 3
f3 <- function(x) {1/(1-x)} 
Result3 <- taylor(f3, x0=a3, 5)
Result3

plot(Result3)

# result-1: range of 3 to 5 is a useful linear range

```


** Question-2 **

2). 
f(x) = e^x
if centered at c=0

$$f(x) = e^x$$
$$f(0) = 1$$


Derivatives:

$$f'(0) = e^x = 1$$

$$f''(0) = e^x = 1$$

$$f'''(x) = e^x = 1$$

Taylor Series:
$$\Sigma_{n=0}^{\infty}\frac{f'^{(n)}(0)}{(n!)}(x)^n = e^x + \frac{e^{(x)}}{(1!)}(x)^{1} + \frac{e^{(x)}}{(2!)}(x)^{2} + ...$$

$$  = 1 + \frac{1}{(1!)}(x)^{1} + \frac{1}{(2!)}(x)^{2} + ...$$

when plug in functions equals:

$$  = \Sigma_{n=0}^{\infty}\frac{1}{(n!)}(x)^n $$

Calculate in R:


``````{r, echo=TRUE}
# check range -8 to 8
a21 = -8
f21 <- function(x) exp(x)
Result21 <- taylor(f21, x0=a21, 8)
Result21

plot(Result21)

# check range 0 to 8
a22 = 0
f22 <- function(x) exp(x)
Result22 <- taylor(f22, x0=a22, 8)
Result22

plot(Result22)

# check range 0 to 4
a23 = 0
f23 <- function(x) exp(x)
Result23 <- taylor(f23, x0=a23, 4)
Result23

plot(Result23)

# Result-2: range of 0 to 4 is a useful linear range



```

** Question-3 **

3). 
f(x) = ln(1+x)
if centered at 0
$$f(x) = ln(1+x)$$


Derivatives:
$$f'(0) = \frac{1}{(1+x)} = 1$$


$$f'"(0) = \frac{1}{(1+x)^2} = -1$$


$$f'''(x) = \frac{2}{(1+x)^3} = 2$$
$$\frac{d^4}{d(x)^4}ln(1+x) = \frac{-6}{(1+x)^4} = -6$$
$$\frac{d^5}{d(x)^5}ln(1+x) = \frac{24}{(1+x)^5} = 24$$


Taylor Series:
$$\Sigma_{n=0}^{\infty}\frac{f'^{(n)}(0)}{(n!)}(x)^n = ln(1+x) + \frac{1}{(1!)}(x)^{1} + \frac{-1}{(2!)}(x)^{2} + ...$$
$$ = 0 + \frac{1}{(1!)}(x)^{1} + \frac{-1}{(2!)}(x)^{2} + \frac{3}{(3!)}(x)^{3} ...$$

$$ = x - \frac{1}{(2)}(x)^{2} + \frac{1}{(3)}(x)^{3} - \frac{1}{(4)}(x)^{4} ...$$
$$ = \Sigma_{n=0}^{\infty} (-1)^{n+1} \frac{1}{(n)}(x)^n $$



Calculate in R:



``````{r, echo=TRUE}

# check range 0 to 8
a31 = 0
f31 <- function(x) log((1 + x), base = exp(1))
Result31 <- taylor(f31, x0=a31, 8)
Result31

plot(Result31)

# check range 0 to 6
a32 = 0
f32 <- function(x) log((1 + x), base = exp(1))
Result32 <- taylor(f32, x0=a22, 6)
Result32

plot(Result32)

# check range 0 to 3
a33 = 0
f33 <- function(x) log((1 + x), base = exp(1))
Result33 <- taylor(f33, x0=a22, 3)
Result33

plot(Result33)

# Result-3: range of 0 to 8 is a better representaive range with this function.  It can be seen there is a diverging oscillating output. 


```


** END **


``````{r, echo=TRUE}

```


